---
title: "exercise_week5"
author: "Nicolas Steinmann"
date: "28 5 2021"
output: html_document
---

```{r}
library(readr)
library(sf)
library(terra)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tmap)

wildschwein_BE <- read_delim("data/wildschwein_BE_2056.txt", ",")%>%
  st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE)
```

## Task 1

1) The dataset contains information about which crops are grown on the different fields in our study area
2) The data type is Polygons.
3) There is a unique identifier field which is an integer. For the geometry each field contains a list with a vector of integer coordinates that define the postition and geometry of the feature and an attribute that tells the sf-function which geometry type to use.
4) Judging from the coordinate format this is in the Swiss metric coordinate system LV95.

```{r}
fields <- read_sf("data/Feldaufnahmen_Fanel.gpkg")

ggplot(data = fields)+
  geom_sf(aes(fill = Frucht))+
  coord_sf(datum = 2056)
```

## Task 2

```{r}
#select may and june fixes
wildschwein_may_june <- filter(wildschwein_BE, month(wildschwein_BE$DatetimeUTC) == 5 | month(wildschwein_BE$DatetimeUTC) == 6)

#Spatial join
joined_wildschwein <- st_join(wildschwein_may_june, fields, st_intersects)

```

## Task 3


Ruth loves Onions and Corn
Sabi is into beans and wetlands
Rosa likes to eat barley and beets.

Everyones chillin' in the forest most of the time.
```{r}

grouped_wildschwein <- joined_wildschwein%>%
  group_by(Frucht, TierName)%>%
  count(Frucht)%>%
  ungroup()%>%
  
  group_by(Frucht)%>%
  mutate(ntot = sum(n), nrel = (n/ntot)*100)%>%
  ungroup()%>%
  
  filter(ntot > 300)


ggplot(grouped_wildschwein)+
  geom_bar(aes(x = Frucht, y = nrel, fill = TierName), stat = 'identity')+
  labs(title = "Relative Verteilung der Aufenthaltsorte")+
  ylab("[%]")+
  xlab("Feldfrucht")


ggplot(grouped_wildschwein)+
  geom_bar(aes(x = Frucht, y = n, fill = TierName), stat = 'identity')+
  labs(title = "Absolute Verteilung der Aufenthaltsorte")+
  ylab("Anzahl Fixe")+
  xlab("Feldfrucht")
  


#Preparing temporal and relative values
grouped_wildschwein2 <- joined_wildschwein%>%
  mutate(hour = hour(DatetimeUTC))%>%
  group_by(TierName, hour)%>%
  add_tally(name = "ntot")%>%
  group_by(TierName, hour, Frucht, ntot)%>%
  count()%>%
  mutate(nrel = round((n/ntot)*100, 2))%>%
  
  #Preparing data so it has less irrelevant classes
  mutate(Frucht = ifelse(is.na(Frucht)==TRUE, "Andere", Frucht), newFrucht = ifelse(nrel >= 5, Frucht, "Andere"))
  
  

ggplot(grouped_wildschwein2)+
  geom_bar(width = 1, aes(x = hour, y = nrel, fill = newFrucht), stat = 'identity')+
  facet_wrap(~TierName, nrow = 1)+
  labs(title = "Aufenthaltsorte von Wildschweinen über den Tagesverlauf (Relativ)")+
  ylab("[%]")+
  xlab("Stunde")


ggplot(grouped_wildschwein2)+
  geom_bar(width = 1, aes(x = hour, y = n, fill = newFrucht), stat = 'identity')+
  facet_wrap(~TierName, nrow = 1)+
  labs(title = "Aufenthaltsorte von Wildschweinen über den Tagesverlauf (Absolute)")+
  ylab("Anzahl Fixes")+
  xlab("Stunde")


```
##Task 4

```{r}
veg_index <- terra::rast("data/vegetationshoehe_LFI.tif")

#plotting the raster data
tmap_mode("view")

tm_shape(veg_index, scientfic = TRUE, format = "g") +
  tm_raster("vegetationshoehe_LFI", palette = viridisLite::viridis(10))

```
##Task 5

Boars come out of the forest in the evening and go back in the morning.
```{r}
#creating a vector
xy <- joined_wildschwein
st_geometry(xy) <- NULL
xy <- vect(xy, type="points", geom=c("E", "N"))

#extracting the values from the raster
veg_info <- terra::extract(veg_index, xy, method = "simple", xy = TRUE, fun = NULL)

#joining the annotated coordinates to the spatial object with a left join
annotated_wildschwein <- left_join(joined_wildschwein, veg_info, by = c("E" = "x", "N" = "y"))


grouped_wildschwein3 <- annotated_wildschwein%>%
  mutate(hour = hour(DatetimeUTC))%>%
  group_by(TierName, hour)%>%
  summarise(veg_med = median(vegetationshoehe_LFI, na.rm = TRUE),
            veg_mean = mean(vegetationshoehe_LFI, na.rm = TRUE),
            veg_std = sd(vegetationshoehe_LFI, na.rm = TRUE)
            )

ggplot(grouped_wildschwein3)+
  geom_bar(width = 1, aes(x = hour, y = veg_mean, fill = TierName), stat = 'identity')+
  facet_wrap(~TierName, nrow = 1)+
  labs(title = "Durchschnittliche Vegetationshöhe beim Aufenthaltsort von Wildschweinen")+
  ylab("Durchschnittlich Vegetationshöhe [m]")+
  xlab("Stunde")

ggplot(grouped_wildschwein3)+
  geom_bar(width = 1, aes(x = hour, y = veg_med, fill = TierName), stat = 'identity')+
  facet_wrap(~TierName, nrow = 1)+
  labs(title = "Mediane Vegetationshöhe beim Aufenthaltsort von Wildschweinen")+
  ylab("Mediane Vegetationshöhe [m]")+
  xlab("Stunde")

ggplot(grouped_wildschwein3)+
  geom_bar(width = 1, aes(x = hour, y = veg_std, fill = TierName), stat = 'identity')+
  facet_wrap(~TierName, nrow = 1)+
  labs(title = "Standardabweichung der Vegetationshöhe beim Aufenthaltsort von Wildschweinen")+
  ylab("Sd Vegetationshöhe [m]")+
  xlab("Stunde")


```